 

### 2020-09-10 vertical-thermal-review.Rmd
worked on references
@NidhiVinod , I switched to using Zotero_references.bib. I also regenerated it from Zotero because for some reason it had a different convention for article names.

---
title: ''
output: word_document
bibliography: references.bib #This can be generated by exporting BibTex file from Zotero
csl: new-phytologist.csl

@@

**Wind speeds are also higher at the top of the canopy (Fig. **2x**).**
*(fill in some specifics from the NEON analysis)*
Consistent with the results from NEON sites, higher wind speed at greater heights within a forest have previously been observed in both closed canopy forests [e.g., REFS; @mcgregor_tree_nodate] and open savannas (Curtis et al. 2018).
This results in higher boundary layer conductance, *DEFINE*, for canopy leaves [@martin_boundary_1999; Roberts et al. 1990].

**Air temperature, $T_{air}$, is sometimes significantly buffered by forest canopies (Fig. **2x**).**
Studies comparing $T_{air}$ under forest canopies with nearby clearings have found lower maximum temperatures under forest canopies across Europe [@zellweger_seasonal_2019] and in the northwestern United States [@davis_microclimatic_2019]. 
Similarly, maximum air temperatures were higher above than below tropical forest canopies during wet seasons in Panama (Rey-Sánchez et al. 2016) and coastal Brazil (Fauset et al. 2018).
However, similar maximum temperatures have been observed during the dry season in Panama (Rey-Sánchez et al. 2016) and in a temperate deciduous forest in the eastern United States [@mcgregor_tree_nodate].
Minimum $T_{air}$ is also buffered by forest canopies under some conditions. 
Higher minimum temperatures under forest canopies relative to nearby clearings have been observed across Europe [@zellweger_seasonal_2019] and in the northwestern United States [@davis_microclimatic_2019].
However, similar minimum temperatures under forest canopies have been observed in tropical forests in Panama (Rey-Sánchez et al. 2016) and coastal Brazil (Fauset et al. 2018), as well as in a temperate deciduous forest in the eastern United States [@mcgregor_tree_nodate].
*(Add something about risk of freeze.)*

**Humidity also varies across the forest vertical profile, being generally higher in the understory (Fig. **2x**).**
- DETAILS - REFS
- @mcgregor_tree_nodate
RH, in combination with temperature, determines vapor pressure deficit, $VPD$
- $VPD$ is lower in understory of Atlantic forest during wet season (Fauset et al. 2018)
This means that canopy leaves tend to be exposed to higher *evaporative demand*.

@@

**The strength of this buffering varies across forests, being influenced by both forest characteristics and the biophysical environment.**
First, buffering increases with canopy cover.
- greater cover --> lower max T and VPD, higher minT  [@davis_microclimatic_2019]
- greater cover -->  lower max T [@zellweger_seasonal_2019]
- [@thom_effects_2020]
*Presumably, buffering would also be affected by canopy roughness, which affects turbulent air flow and the canopy boundary layer.*
Taller trees don’t necessarily increase buffering [@zellweger_seasonal_2019]. 
SCA species increase T buffering (Zelllweger et al. 2019) [@zellweger_seasonal_2019]
The strength of buffering also varies with respect to geographic and climatic factors.
- Distance to coast, topographic position, elevation [@zellweger_seasonal_2019]
- [@davis_microclimatic_2019]

@@

```{r eval=FALSE, echo=FALSE}
# refs in this table need to go in the nocite field below
```
---
nocite: | 
  @zellweger_seasonal_2019 
---

@@

*Max stomatal conductance*

@@

![(*For latest version of table, see tables.pdf.*)](table2.png)
```{r eval=FALSE, echo=FALSE}
# refs in this table need to go in the nocite field below
```
---
nocite: | 
  @zellweger_seasonal_2019 
---

>>>>

---
title: ''
output: word_document
bibliography: Zotero_references.bib #This can be generated by exporting BibTex file from Zotero
csl: new-phytologist.csl

@@

**Wind speeds are also higher at the top of the canopy (Fig. **2x**).**
*(fill in some specifics from the NEON analysis)*
Consistent with the results from NEON sites, higher wind speed at greater heights within a forest have previously been observed in both closed canopy forests [e.g., REFS; @mcgregor_tree_nodate] and open savannas (Curtis et al. 2018).
This results in higher boundary layer conductance, *DEFINE*, for canopy leaves [@martin_boundary_1999; @roberts_stomatal_1990].

**Air temperature, $T_{air}$, is sometimes significantly buffered by forest canopies (Fig. **2x**).**
Studies comparing $T_{air}$ under forest canopies with nearby clearings have found lower maximum temperatures under forest canopies across Europe [@zellweger_seasonal_2019] and in the northwestern United States [@davis_microclimatic_2019]. 
Similarly, maximum air temperatures were higher above than below tropical forest canopies during wet seasons in Panama [@rey-sanchez_spatial_2016] and coastal Brazil [@fauset_differences_2018].
However, similar maximum temperatures have been observed during the dry season in Panama [@rey-sanchez_spatial_2016] and in a temperate deciduous forest in the eastern United States [@mcgregor_tree_nodate].
Minimum $T_{air}$ is also buffered by forest canopies under some conditions. 
Higher minimum temperatures under forest canopies relative to nearby clearings have been observed across Europe [@zellweger_seasonal_2019] and in the northwestern United States [@davis_microclimatic_2019].
However, similar minimum temperatures under forest canopies have been observed in tropical forests in Panama [@rey-sanchez_spatial_2016] and coastal Brazil [@fauset_differences_2018], as well as in a temperate deciduous forest in the eastern United States [@mcgregor_tree_nodate].
*(Add something about risk of freeze.)*

**Humidity also varies across the forest vertical profile, being generally higher in the understory (Fig. **2x**).**
- DETAILS - REFS
- @mcgregor_tree_nodate
RH, in combination with temperature, determines vapor pressure deficit, $VPD$
- $VPD$ is lower in understory of Atlantic forest during wet season [@fauset_differences_2018]
This means that canopy leaves tend to be exposed to higher *evaporative demand*.

@@

**The strength of this buffering varies across forests, being influenced by both forest characteristics and the biophysical environment.**
First, buffering increases with canopy cover.
* greater cover --> lower max T and VPD, higher minT  [@davis_microclimatic_2019]
* greater cover -->  lower max T [@zellweger_seasonal_2019]
* [@thom_effects_2020]
*Presumably, buffering would also be affected by canopy roughness, which affects turbulent air flow and the canopy boundary layer.*
Taller trees don’t necessarily increase buffering [@zellweger_seasonal_2019]. 
SCA species increase T buffering (Zelllweger et al. 2019) [@zellweger_seasonal_2019]
The strength of buffering also varies with respect to geographic and climatic factors.
* Distance to coast, topographic position, elevation [@zellweger_seasonal_2019]
* [@davis_microclimatic_2019]

@@

```{r eval=FALSE, echo=FALSE}
# refs in this table need to go in the nocite field below (easiest in alphabetical order)
```
---
nocite: | 
  @mau_temperate_2018
---

@@

@@

![(*For latest version of table, see tables.pdf.*)](table2.png)
```{r eval=FALSE, echo=FALSE}
# refs in this table need to go in the nocite field below (easiest in alphabetical order)
```
---
nocite: | 
  @bolstad_foliar_1999; @roberts_stomatal_1990; @slot_photosynthetic_2019 
---
